---
title: "Game Collecting Analysis"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  highlight: kate
toc_depth: 6
---

```{r knitr_init, cache = FALSE, warning = FALSE, echo = FALSE}
library(knitr)
## Global options
opts_chunk$set(echo = FALSE,
               cache = FALSE,
               prompt = FALSE,
               comment = NA,
               include = TRUE,
               message = FALSE,
               warning = FALSE)
```

```{r globals}
# Data comes from a survey of a small game collecting community
DATA_PATH <- file.path(dirname(getwd()), "data", "survey-data-2020-05-06.csv")

RESPONSE_SEP <- ";"

# Anything with "likelihood" is on a 1-5 scale with 1 being very likely and 5 being very unlikely
# Columns that allow multiple responses separated by ";"
#   system_types
#   console_generations
#   handheld_generations
#   thought_process
# Other columns contain single responses and can be treated as factors
COLUMN_RECODE_NAMES <- c(
  "timestamp",
  "system_types",
  "console_generations",
  "handheld_generations",
  "before_your_time",
  "current_systems",
  "system_variant_likelihood",
  "sealed_likelihood",
  "cib",
  "missing_inserts",
  "loose",
  "wear_likelihood",
  "illegible_likelihood",
  "price_likelihood",
  "wishlist",
  "system_complete",
  "series_complete",
  "price_check",
  "never_played_before",
  "review_influence_likelihood",
  "research_likelihood",
  "thought_process",
  "age",
  "gender",
  "ethnicity",
  "education",
  "employment"
)

GENERATION_LEVELS <- c(
  "First",
  "Second",
  "Third",
  "Fourth",
  "Fifth",
  "Sixth",
  "Seventh",
  "Eighth",
  "Ninth"
)

CIB_LEVELS <- c(
  "I only collect CIB games.",
  "It varies by system.",
  "It varies by game series.",
  "It varies by cost of CIB.",
  "I don't specifically seek out CIB games."
)
```

```{r packages}
library(dplyr)
library(DT)
library(plotly)
library(UpSetR)
```

```{r functions}
# Function to turn a vector into a dataframe of counts with known column names c("categories", "count")
# Input:
#   some_vector = a vector of values to be tabluated (string | factor recommended)
# Output:
#   a data.frame with a two columns: categories as a factor and count as a numeric
df_from_vector <- function(some_vector, factor_levels=c()) {
  df <- as.data.frame(table(some_vector, dnn = list("categories")), responseName = "count")
  factor_levels <- if (length(factor_levels) == 0) { unique(df$categories) } else { factor_levels }
  df$categories <- factor(
    df$categories,
    levels = factor_levels
  )
  df
}

# Function to turn a multiresponse vector into an observation matrix for plotting with UpSetR
# Input:
#   multiresponse_vector is a vector of multiple response values each set of multiple responses
#     is separated by RESPONSE_SEP
#   RESPONSE_SEP is the string separator of multiple responses
# Output:
#   a dataframe with one column for "Respondents" and additional columns for each unique response
#     the values in the unique response columns are either 1 (checked) or 0 (not checked)
upset_format <- function(multiresponse_vector, response_sep = RESPONSE_SEP) {
  split_responses <- strsplit(as.character(multiresponse_vector), response_sep)
  unique_responses <- unique(unlist(split_responses))
  response_df <- bind_rows(
    lapply(
      split_responses,
      function(split_response) {
        sapply(
          unique_responses,
          function(known_response) {if (known_response %in% split_response) {1} else {0}}
        )
      }
    )
  )
  cbind(data.frame(respondent = paste0("Respondent_", seq(nrow(response_df)))), response_df)
}


## TODO: Create a function for generating horizontal bar charts from count data.frames

## TODO: Create a function for generating likelihood plots
```

```{r read-n-format-data}
survey_df <- read.csv(DATA_PATH)
colnames(survey_df) <- COLUMN_RECODE_NAMES
```

# Introduction

This project is a simple graphical exploration of results from a survey of a small game collecting group.
There were a total of `r nrow(survey_df)` respondents.

# Which systems do you collect for?

## Figure 1, System Types

```{r overarching-system-types, fig.height=5, fig.width=8}
upset(
  data = upset_format(survey_df$system_types),
  empty.intersections = 2,
  
  mb.ratio = c(0.5, 0.5),
  main.bar.color = "black",
  matrix.color = "black",
  sets.bar.color = "black",
  text.scale = 1.7
)
```

This is an [UpSet](http://gehlenborglab.org/research/projects/upsetr/) plot.
It acts as an alternative to a venn diagram.
You can see the overall number of respondents collecting for each system type on the left.
To the right, you see how many respondents collect for each combination of systems.
`

The overwhelming majority of respondents collect for consoles and handhelds.
No respondent collects for just PC or just handhelds and anyone that collects for PC also collects for consoles and handhelds.

## Figure 2, Console Generations
```{r console-generations}
upset(
  data = upset_format(gsub(" Generation \\([[:alnum:][:space:],/-]+)", "", survey_df$console_generations)),
  mb.ratio = c(0.5, 0.5),
  main.bar.color = "black",
  matrix.color = "black",
  sets.bar.color = "black",
  text.scale = 1.7,
  nsets = 20
)
```

```{r console-generations-table}
console_generations <- unique(unlist(strsplit(as.character(survey_df$console_generations), RESPONSE_SEP)))
console_generations_df <- data.frame(
  Generation = factor(gsub(" Generation.*", "", console_generations), levels = GENERATION_LEVELS),
  Consoles = gsub(".*\\(|)", "", console_generations)
)
DT::datatable(
  console_generations_df[order(console_generations_df$Generation),],
  class = 'cell-border stripe hover',
  rownames = FALSE,
  extensions = list("Buttons" = NULL),
  options = list(
    buttons = list(
      c(extend = "copy", text = "Copy Table"),
      "csv",
      "excel",
      c(extend = "colvis", text = "Hide/Show Columns")
    )
  )
)

```

This is another UpSet plot looking at which console generations respondents collect.


Excluding the newest (and still hard to find) consoles, the recent generations are the most popular to collect for.
Overall, the most popular generation is the sixth (PS2, Gamecube, Xbox, Dreamcast).
Only two respondents collect for a single generation and only three respondents collect for just two generations.

## Figure 3, Handheld Generations
```{r handheld-generations}
upset(
  data = upset_format(
    gsub(
      " Generation \\([[:alnum:][:space:].&,/-]+)",
      "",
      subset(survey_df, handheld_generations != "None of the above")$handheld_generations
    )
  ),
  mb.ratio = c(0.5, 0.5),
  main.bar.color = "black",
  matrix.color = "black",
  sets.bar.color = "black",
  text.scale = 1.7,
  nsets = 20
)
```

```{r handheld-generations-table}
handheld_generations <- unique(unlist(strsplit(as.character(survey_df$handheld_generations), RESPONSE_SEP)))
handheld_generations_df <- data.frame(
  Generation = factor(gsub(" Generation.*", "", handheld_generations), levels = GENERATION_LEVELS),
  Consoles = gsub(".*\\(|)", "", handheld_generations)
)
DT::datatable(
  handheld_generations_df[order(handheld_generations_df$Generation),],
  class = 'cell-border stripe hover',
  rownames = FALSE,
  extensions = list("Buttons" = NULL),
  options = list(
    buttons = list(
      c(extend = "copy", text = "Copy Table"),
      "csv",
      "excel",
      c(extend = "colvis", text = "Hide/Show Columns")
    )
  )
)

```

Most everyone that collects handhelds collects for the same generations (Fourth, Sixth, Seventh, and Eight).

## Figure 4, Ease of access

```{r retro-bars, fig.height=3, fig.width=8}
retro_ly <- plot_ly(
  data = df_from_vector(survey_df$before_your_time),
  x = ~categories,
  y = ~count,
  type = "bar"
) %>%
  layout(
    yaxis = list(title = "Count"),
    xaxis = list(title = ""),
    name = "Retro"
  )

new_ly <- plot_ly(
  data = df_from_vector(survey_df$current_systems),
  x = ~categories,
  y = ~count,
  type = "bar"
) %>%
  layout(
    yaxis = list(title = "Count"),
    xaxis = list(title = ""),
    name = "New"
  )


subplot(retro_ly, new_ly, shareY = TRUE) %>%
  layout(annotations = list(
    list(
      x = -0.01,
      y = 1.5,
      text = "Do you collect systems you did not own/have\neasy access to before they were discontinued?",
      showarrow = F,
      xref='paper',
      yref='paper'
    ),
  list(
    x = 0.9,
    y = 1.5,
    text = "Do you collect systems\ncurrently being manufactured?",
    showarrow = F,
    xref='paper', 
    yref='paper')
  ),
  margin = 0,
  showlegend = FALSE
)
```

A majority of respondents collect for at least one system they did not own or have easy access to before those systems were dicsontinued.
A majority of respondents also collect for currently manufactured systems.

## Figure 5, Collecting multiple system variants
```{r hardware-variant-collecting, fig.height=3, fig.width=6}
plot_ly(
  data = survey_df,
  x = ~system_variant_likelihood,
  type = "histogram"
) %>%
  layout(
    title = "How likely are you to collect multiple variations of hardware?",
    yaxis = list(title = "Count"),
    xaxis = list(title = "", tickvals = c(1, 5), ticktext = c("Very Likely", "Very Unlikely")),
    bargap = 0.1
  )

```

Despite the abundance of multiple variant hardware posts we see, most respondents are unlikely to purchase system variants.

# What condition of games do you collect?

## Figure 6, Sealed and staying sealed
```{r sealed, fig.height=4, fig.width=5, fig.align='right'}
plot_ly(
  data = df_from_vector(survey_df$sealed_likelihood, c("Never", "Sometimes", "Frequently")),
  y = ~categories,
  x = ~count,
  type = "bar",
  orientation = 'h'
) %>% 
  layout(
    title = "How often do you collect factory sealed games\nwithout the intention of opening them?",
    yaxis = list(title = ""),
    xaxis = list(title = "Count"),
    margin = 0
  )
```

Perhaps unsurprisingly, most respondents do not often collect games without the intention of opening them.

## Figure 7, Prioritizing CIB
```{r cib, fig.height=4, fig.width=8}
plot_ly(
  data = df_from_vector(survey_df$cib, CIB_LEVELS),
  y = ~categories,
  x = ~count,
  type = "bar",
  orientation = 'h'
) %>%
  layout(
    title = "Do you collect complete in box (CIB) games?",
    yaxis = list(title = ""),
    xaxis = list(title = "Count"),
    margin = 0
  )
```

The two most important factors when deciding on collecting CIB or not are the system and the price.

## Figure 8, Missing inserts
```{r missing, fig.height=4, fig.width=8}
plot_ly(
  data = df_from_vector(gsub("cases ", "cases\n", survey_df$missing_inserts)),
  y = ~categories,
  x = ~count,
  type = "bar",
  orientation = 'h'
) %>%
  layout(
    title = "Do you collect games with boxes/cases that\nare missing manuals or other inserts?",
    yaxis = list(title = ""),
    xaxis = list(title = "Count"),
    margin = 0
  )

```

Most respondents prioritize games with cases but don't care if some of the inserts are missing.

## Figure 9, Loose games
```{r loose, fig.height=4, fig.width=8}
plot_ly(
  data = df_from_vector(survey_df$loose),
  y = ~categories,
  x = ~count,
  type = "bar",
  orientation = 'h'
) %>%
  layout(
    title = "Do you collect loose games?",
    yaxis = list(title = ""),
    xaxis = list(title = "Count"),
    margin = 0
  )
```

Most respondents will also collect loose games for certain systems.
In a follow up survey it may be of interest to split this by cartridge and disc based systems.

## Figure 10, Visible wear
```{r wear, fig.height=3, fig.width=6}
plot_ly(
  data = survey_df,
  x = ~wear_likelihood,
  type = "histogram"
) %>%
  layout(
    title = "How likely is it that you would purchase\na game with visible signs of wear?",
    yaxis = list(title = "Count"),
    xaxis = list(title = "", tickvals = c(1, 5), ticktext = c("Very Likely", "Very Unlikely")),
    bargap = 0.1,
    margin = 0
  )
```

On average, respondents are ambivalent towards moderate wear and tear.
This could be indicate that minor wear and tear is not as important as other factors such as price.

## Figure 11, Superficial Damage
```{r damage, fig.height=3, fig.width=6}
plot_ly(
  data = survey_df,
  x = ~illegible_likelihood,
  type = "histogram"
) %>%
  layout(
    title = "How likely is it that you would purchase a game with a\nheavily damaged and illegible or missing label?",
    yaxis = list(title = "Count"),
    xaxis = list(title = "", tickvals = c(1, 5), ticktext = c("Very Likely", "Very Unlikely")),
    bargap = 0.1,
    margin = 0
  )
```

Most respondents are extremely unlikely to purchase a game with a significant damage.

## Figure 12, Price vs CIB
```{r price, fig.height=3, fig.width=6}
plot_ly(
  data = survey_df,
  x = ~price_likelihood,
  type = "histogram"
) %>%
  layout(
    title = "How likely is price to influence your decision to purchase\na sealed, CIB, case only, or loose game?",
    yaxis = list(title = "Count"),
    xaxis = list(title = "", tickvals = c(1, 5), ticktext = c("Very Likely", "Very Unlikely")),
    bargap = 0.1,
    margin = 0
  )
```

Price seems to be a major driver behind respondents choice to purchase a CIB game.